Redshift Evolution of the Underlying Type Ia Supernova Stretch Distribution

### 2020-08-18 Nicolas.tex
New image

\newcommand{\nn}[1]{\textcolor[rgb]{0.25, 0.50, 0}{#1}}
\newcommand{\prob}[2]{\mathcal{P}\left( #1 \mid #2\right)}

@@

\subsection{\nn{Verifying completeness}}\label{ssec:verify}

\nn{In order to ensure that the effects we model are only due to the supposed drift,
we need to check for possible Malmquist bias left in the sample. Because color
is most sensitive to selection effects, we studied both the stretch and color of
the end of SDSS and the start of PS1, for $0.10 < z < 0.20$. In this redshift
range, because the SDSS SNe between $0.15 < z < 0.20$ are between our
conservative and fiducial cuts, these SNe are the most prone to have selection
effects. However, the PS1 dataset is quite robust because these SNe are far from
both the conservative and fiducial cuts; see Fig. \ref{fig:cuts}.}

\nn{The results are shown Fig. \ref{fig:distrib}. We represented the normed
histograms along with their error bars, and found that they don't differ much.
To ascertain this idea, we used a Kolmogorov-Smirnov test that gave us a p-value
of $0.265$ for the color, and $0.137$ for the stretch; while it doesn't tell us
that both the SDSS and PS1 parts come from the same distribution, as would be
expected if we were free from selection effects (so that the samples were random
draws from what Nature could give us), it is not excluded. }

\begin{figure*}
    \centering
    \includegraphics[width=17cm]{Article_figures/both-cut_SDSS_PS1-010-020.pdf}
    \caption{\nn{$x_1$ and $c$ normed histograms of the SDSS and PS1 samples for
        $0.10 < z < 0.20$ and their errorbars. On the left, we find in orange
        lines the Base model distributions for $z = 0.10, 0.15, 0.20$ from top
        to bottom. The p-values are the results from a Kolmogorov-Smirnov test;
    they don't show any indication that the samples are not taken from the same
distribution, for both $c$ and $x_1$.}}
    \label{fig:distrib}
\end{figure*}

@@

>>>>

\newcommand{\nn}[1]{\textcolor[rgb]{0.25, 0.50, 0}{#1}}
\newcommand{\mr}[1]{\textcolor[rgb]{0.5, 0.20, 0.3}{#1}}
\newcommand{\prob}[2]{\mathcal{P}\left( #1 \mid #2\right)}

@@

\subsection{\mr{Testing the construction of a volume limited sample}}\label{ssec:verify}

\mr{In this analysis, we have built a volume limited sample from a set of
magnitude limited ones, using a simple redshift limit.  A ongoing follow up
analysis will make use of the SNANA package \citep{SNANA} to account for
more subtle selection functions. However, the strength of the current
approach is that it is independent of potential inaccuracy in the modeling
of the complex selection functions of the aforementioned dataset. Yet, it is
necessary to test if the sample we have created could show sign of leftover
sample selections.}

\mr{To do so, we compare the stretch and color distributions of the SNe~Ia in
our sample between datasets of overlapping redshifts. If indeed we have built
a volume limited sample, the distributions of SNe~Ia parameters at a given
redshift should not depend of their origin surveys. When considering a
redshift range, this has to be narrow enough such that any drift would be
negligible. The two samples that overlap the most in redshift are PS1 and
SDSS in the redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}).
This corresponds to the 146 SNe~Ia from SDSS at the highest redshift end and
thus potentially the most affected by potential selection effect leftover
(see the corresponding discussion in section~\ref{sec:sample}). In that same
redshift range, PS1 has 52 SNe~Ia, which are at in the lowest redshift bin
and thus unlikely to have any selection function issues. We present the
stretch and color distributions of both samples in the redshift range $0.10 <
z < 0.20$ in Fig.~\ref{fig:distrib}, where we also provide the p-values of
Kolmogorov-Smirnov tests. The high p-values both in color and stretch
distributions ($>10\%$) strengthen the visual impression that there is no
reason to believe that the SDSS and PS1 SNe~Ia in their overlapping redshift
range are not drawn from the same underlying distributions. A similar
analysis is doable for PS1 and SNLS in their redshift range $0.20 < z <
0.30$ where the same conclusion could be drawn, i.e., that there is no sign
of discrepancy between in the stretch and color distributions between the
low- and high-end of the SNLS and PS1 datasets, respectively. Nonetheless,
the small size of the SNLS dataset at $z < 0.30$ (12 SNe~Ia vs. 90 for PS1)
limits the interest of the test as only strong deviation would be
noticeable. Extending the redshift range to $0.20 < z < 0.40$ (though we have no
PS1 data above 0.3) enables to get 31 SNe~Ia from SNLS in total, the stretch
and color p-values remains high, at 34\% and 4\% respectively.}

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/both-cut_SDSS_PS1-010-020.pdf}
    \caption{\nn{$x_1$ and $c$ normed histograms of the SDSS and PS1 samples for
        $0.10 < z < 0.20$ and their errorbars. On the left, we find in orange
        lines the Base model distributions for $z = 0.10, 0.15, 0.20$ from top
        to bottom. The p-values are the results from a Kolmogorov-Smirnov test;
    they don't show any indication that the samples are not taken from the same
distribution, for both $c$ and $x_1$.}}
    \label{fig:distrib}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/both-cut_SNLS_PS1-020-030.pdf}
    \caption{\mr{Same as Fig.~\ref{fig:distrib} for the PS1 and SNLS data in
    their overlapping redshift range $0.20 < z < 0.30$.}}

    \label{fig:distrib-ps-snls}
\end{figure}

\mr{We highlight that the SNe~Ia color is more prone to selection effects than
stretch as illustrated in Fig.~\ref{fig:maglim}; see also e.g., Fig.~3 of
\cite{kessler2017}. Hence, since the color shows no sign of leftover selection
effects, it further supports our claim that our simple redshift-based selection
criteria is sufficient to test the redshift drift of the SN~Ia stretch
distribution.}

\nn{In order to ensure that the effects we model are only due to the supposed
drift, we need to check for possible Malmquist bias left in the sample.
Because color is most sensitive to selection effects, we studied both the
stretch and color of the end of SDSS and the start of PS1, for $0.10 < z <
0.20$. In this redshift range, }

\nn{The results are shown Fig. \ref{fig:distrib}. We represented the normed
histograms along with their error bars, and found that they don't differ much.
To ascertain this idea, we used a Kolmogorov-Smirnov test that gave us a p-value
of $0.265$ for the color, and $0.137$ for the stretch; while it doesn't tell us
that both the SDSS and PS1 parts come from the same distribution, as would be
expected if we were free from selection effects (so that the samples were random
draws from what Nature could give us), it is not excluded.}

@@

\bibitem[Kessler et al.(2009)]{SNANA} Kessler, R., Bernstein, J.~P., Cinabro,
D., et al.\ 2009, \pasp, 121, 1028
